---
title: "Networks Project"
output: html_document
date: "2023-06-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load Libraries.
```{r}
library(incidentally)     # install.packages("incidentally")
library(backbone)         # install.packages("backbone")
library(igraph)
library(tibble)
library(stringr)
library(dplyr)
library(tidyr)
```


## SENATE - Pull incidence matrices and data for each session 
```{r}
# Create an empty list to store the outputs
senate_incidences <- list()

# Define the range of sessions (years)
start_session <- 108
end_session <- 118

# Loop through each session and extract the incidence matrix
for (session in start_session:end_session) {
  # Initialize an empty variable to store the current session's output
  senate_session <- NULL
  
  tryCatch({
  # Extract incidence matrix for the current session
  senate_session <- incidence.from.congress(
    session = session,
    format = "data",
    types = c("s", "sjres")
  ) 
  
  # Generate a name for the output list item
  session_name <- paste0("senate_", session)
  
  # Save the incidence matrix to the output list
  senate_incidences[[session_name]] <- senate_session
  
  # Print a message indicating the progress
  cat("Session", session, "completed.\n")
  }, error = function(e) {
    # Print an error message if the extraction fails
    cat("Error extracting session", session, ": ", conditionMessage(e), "\n")
  })
}


# Print the names of the output list items
print(names(senate_incidences))

```



SENATE - Extract backbone using SDSM and save
```{r}

senate_list <- list()
  
for (i in seq_along(senate_incidences)) {
  tryCatch({
  # Save name to remember 
  name <- names(senate_incidences)[i]
    
  # First extract the matrix from the list extracted from incidentally
  object <- senate_incidences[[i]]
  
  # Extract backbone using the Stochastic Degree Sequence Model
  matrix <- as.matrix(sdsm(object$matrix, narrative = TRUE, signed = TRUE))
  
  # Then create the dataframe
  df <- createDataFrame(matrix)
  
  # Then create the legislators lists
  dems_list <- subset(object$legislator, party == "D")
  reps_list <- subset(object$legislator, party == "R")

  # Append to senate_list
  senate_list[[i]] <- tibble::lst(name, object, matrix, df, dems_list, reps_list)
  }, error = function(e) {
    # Print an error message if the extraction fails
    cat("Error extracting", name, ": ", conditionMessage(e), "\n")
  })
}

```



```{r}
names(senate_list) <- c(108, 111, 113, 114, 115, 116, 117, 118)
```

Save the output.
```{r}
save(senate_incidences, 
     senate_list, 
     file="senate_data.rdata")
```

*Note: The SDSM function is not working for sessions 109, 110, and 112*
```{r}
senate_109_inc <- senate_incidences$senate_109$matrix
```



## HOUSE - Pull incidence matrices and data for each session
```{r}
# Create an empty list to store the outputs
house_incidences <- list()

# Define the range of sessions (years)
start_session <- 108
end_session <- 118

# Loop through each session and extract the incidence matrix
for (session in start_session:end_session) {
  # Initialize an empty variable to store the current session's output
  house_session <- NULL
  
  tryCatch({
  # Extract incidence matrix for the current session
  house_session <- incidence.from.congress(
    session = session,
    format = "data",
    types = c("hr", "hjres")      # "hjres"    # "hr"
  )
  
  # Generate a name for the output list item
  session_name <- paste0("house_", session)
  
  # Save the incidence matrix to the output list
  house_incidences[[session_name]] <- house_session
  
  # Print a message indicating the progress
  cat("Session", session, "completed.\n")
  }, error = function(e) {
    # Print an error message if the extraction fails
    cat("Error extracting session", session, ": ", conditionMessage(e), "\n")
  })
}


# Print the names of the output list items
print(names(house_incidences))

```


HOUSE - Extract backbone using SDSM and save
```{r}
house_list <- list()
  
for (i in seq_along(house_incidences)) {
  # Save name to remember 
  name <- names(house_incidences)[i]
    
  # First extract the matrix from the list extracted from incidentally
  object <- house_incidences[[i]]
  
  # Extract backbone using the Stochastic Degree Sequence Model
  matrix <- as.matrix(sdsm(object$matrix, narrative = TRUE, signed = TRUE))
  
  # Then create the dataframe
  df <- createDataFrame(matrix)
  
  # Then create the legislators lists
  dems_list <- subset(object$legislator, party == "D")
  reps_list <- subset(object$legislator, party == "R")

  # Append to house_list
  house_list[[i]] <- tibble::lst(name, object, matrix, df, dems_list, reps_list)
}

```

```{r}
names(house_list) <- seq(108, 118)
```

Save the output.
```{r}
save(house_incidences,
     house_list,
     file="house_data.rdata")
```

*Note: The incidence function was failing to pull the House resolution ("hr") bills so the data for House 116th session is only based on 388 bills that were house joint resolution ("hjres") bills*



# Recreating the figures

Function to extract pairs and their tie
```{r}
## Function 
createDataFrame <- function(matrix) {
  # Define upper triangle values
  upper_triangle <- upper.tri(matrix)
  upper_triangle_values <- matrix[upper_triangle]
  
  # Get the row and column names of the upper triangle elements
  rep_pairs <- expand.grid(colnames(matrix), colnames(matrix))
  rep_pairs <- rep_pairs[upper_triangle, ]
  
  # Create the dataframe with representative pairs
  df <- data.frame(rep1 = rep_pairs$Var1,
                   rep2 = rep_pairs$Var2,
                   tie = upper_triangle_values)
  
  # Return the df
  return(df)
  
  # Remove unnecessary stuff
  rm(upper_triangle, upper_triangle_values, rep_pairs, df)
}
```


### FIGURE A - HOUSE
```{r}

prop_pos_all_list_house <- numeric(length(house_list))
prop_neg_all_list_house <- numeric(length(house_list))

for (i in seq_along(house_list)) {
  matrix <- house_list[[i]]$matrix
  prop_pos_all_list_house[i] <- nrow(subset(createDataFrame(matrix), tie == 1)) / nrow(createDataFrame(matrix))
  prop_neg_all_list_house[i] <- nrow(subset(createDataFrame(matrix), tie == -1)) / nrow(createDataFrame(matrix))
}

```

```{r}
# Create df 
figure_A_results <- data.frame (
  session = names(house_list),
  prop_pos_all_H = c(prop_pos_all_list_house),
  prop_neg_all_H = c(prop_neg_all_list_house)
)

figure_A_results
```

#### FIGURE B - SENATE


```{r}

prop_pos_all_list_sen <- numeric(length(senate_list))
prop_neg_all_list_sen<- numeric(length(senate_list))

for (i in seq_along(senate_list)) {
  matrix <- senate_list[[i]]$matrix
  prop_pos_all_list_sen[i] <- nrow(subset(createDataFrame(matrix), tie == 1)) / nrow(createDataFrame(matrix))
  prop_neg_all_list_sen[i] <- nrow(subset(createDataFrame(matrix), tie == -1)) / nrow(createDataFrame(matrix))
}


```

```{r}
# Create df 
figure_B_results <- data.frame (
  session = names(senate_list),
  prop_pos_all_S = c(prop_pos_all_list_sen),
  prop_neg_all_S = c(prop_neg_all_list_sen)
)

figure_B_results
```


```{r}
df<-house_list$`108`$df

nrow(df) == nrow(subset(df, tie==-1)) + nrow(subset(df, tie==1)) + nrow(subset(df, tie==0))
```


### FIGURE C - HOUSE

```{r} 

prop_pos_within_house <- numeric(length(house_list))
prop_neg_between_house <- numeric(length(house_list))

for (i in seq_along(house_list)) {
  # Then create the dataframe
  df <- createDataFrame(house_list[[i]]$matrix)
  
  # Then create the legislators lists
  dems_list <- house_list[[i]]$dems_list
  reps_list <- house_list[[i]]$reps_list
  
  # Add columns "rep1_party" and "rep2_party"
  df$rep1_party <- ifelse(df$rep1 %in% dems_list$name, "D", ifelse(df$rep1 %in% reps_list$name, "R", "NA"))
  df$rep2_party <- ifelse(df$rep2 %in% dems_list$name, "D", ifelse(df$rep2 %in% reps_list$name, "R", "NA"))

  # Append proportions
  prop_pos_within_house[i] <- nrow(subset((subset(df, rep1_party == rep2_party & rep1_party != "NA")), tie == 1)) / nrow(subset(df, tie == 1))
  
  prop_neg_between_house[i] <- nrow(subset((subset(df, rep1_party != rep2_party & rep1_party != "NA" & rep2_party != "NA")), tie == -1)) / nrow(subset(df, tie == -1))
}

```

```{r}
# Create df 
figure_C_results <- data.frame (
  session = names(house_list),
  prop_pos_within_H = c(prop_pos_within_house),
  prop_neg_between_H = c(prop_neg_between_house)
)

figure_C_results
```

#### FIGURE D - SENATE
```{r} 

prop_pos_within_senate <- numeric(length(senate_list))
prop_neg_between_senate <- numeric(length(senate_list))

for (i in seq_along(senate_list)) {
  df <- createDataFrame(senate_list[[i]]$matrix)
  
  # Then create the legislators lists
  dems_list <- senate_list[[i]]$dems_list
  reps_list <- senate_list[[i]]$reps_list
    
  # Add columns "rep1_party" and "rep2_party"
  df$rep1_party <- ifelse(df$rep1 %in% dems_list$name, "D", ifelse(df$rep1 %in% reps_list$name, "R", "NA"))
  df$rep2_party <- ifelse(df$rep2 %in% dems_list$name, "D", ifelse(df$rep2 %in% reps_list$name, "R", "NA"))

  # Append proportions
  prop_pos_within_senate[i] <- nrow(subset((subset(df, rep1_party == rep2_party & rep1_party != "NA")), tie == 1)) / nrow(subset(df, tie == 1))
    
  prop_neg_between_senate[i] <- nrow(subset((subset(df, rep1_party != rep2_party & rep1_party != "NA" & rep2_party != "NA")), tie == -1)) / nrow(subset(df, tie == -1))
}

```

```{r}
# Create df 
figure_D_results <- data.frame (
  session = names(senate_list),
  prop_pos_within_S = c(prop_pos_within_senate),
  prop_neg_between_S = c(prop_neg_between_senate)
)

figure_D_results
```


#### FIGURE E - HOUSE

Get the proportion of negative ties for each party
```{r}

prop_neg_within_dem_house <- numeric(length(house_list))
prop_neg_within_rep_house <- numeric(length(house_list))
  

for (i in seq_along(house_list)) {
  df <- createDataFrame(house_list[[i]]$matrix)
  
  # Then extract the legislators lists
  dems_list <- house_list[[i]]$dems_list
  reps_list <- house_list[[i]]$reps_list
  
  # Add columns "rep1_party" and "rep2_party"
  df$rep1_party <- ifelse(df$rep1 %in% dems_list$name, "D", ifelse(df$rep1 %in% reps_list$name, "R", "NA"))
  df$rep2_party <- ifelse(df$rep2 %in% dems_list$name, "D", ifelse(df$rep2 %in% reps_list$name, "R", "NA"))
  
  # Append proportions
  prop_neg_within_dem_house[i] <- nrow(subset((subset(df, rep1_party == rep2_party & rep1_party == "D")), tie == -1)) / nrow(subset(df, rep1_party == rep2_party & rep1_party == "D"))
    
  prop_neg_within_rep_house[i] <- nrow(subset((subset(df, rep1_party == rep2_party & rep1_party == "R")), tie == -1)) / nrow(subset(df, rep1_party == rep2_party & rep1_party == "R"))
    
}

```

```{r}
# Create df 
figure_E_results <- data.frame (
  session = names(house_list),
  prop_neg_within_dem_H = c(prop_neg_within_dem_house),
  prop_neg_within_rep_H = c(prop_neg_within_rep_house)
)

figure_E_results
```




#### FIGURE F - SENATE

Get the proportion of negative ties for each party
```{r}

prop_neg_within_dem_senate <- numeric(length(senate_list))
prop_neg_within_rep_senate <- numeric(length(senate_list))
  

for (i in seq_along(senate_list)) {
  df <- createDataFrame(senate_list[[i]]$matrix)
  
  # Then extract the legislators lists
  dems_list <- senate_list[[i]]$dems_list
  reps_list <- senate_list[[i]]$reps_list
  
  # Add columns "rep1_party" and "rep2_party"
  df$rep1_party <- ifelse(df$rep1 %in% dems_list$name, "D", ifelse(df$rep1 %in% reps_list$name, "R", "NA"))
  df$rep2_party <- ifelse(df$rep2 %in% dems_list$name, "D", ifelse(df$rep2 %in% reps_list$name, "R", "NA"))
  
  # Append proportions
  prop_neg_within_dem_senate[i] <- nrow(subset((subset(df, rep1_party == rep2_party & rep1_party == "D")), tie == -1)) / nrow(subset(df, rep1_party == rep2_party & rep1_party == "D"))
    
  prop_neg_within_rep_senate[i] <- nrow(subset((subset(df, rep1_party == rep2_party & rep1_party == "R")), tie == -1)) / nrow(subset(df, rep1_party == rep2_party & rep1_party == "R"))
    
}

```

```{r}
# Create df 
figure_F_results <- data.frame (
  session = names(senate_list),
  prop_neg_within_dem_S = c(prop_neg_within_dem_senate),
  prop_neg_within_rep_S = c(prop_neg_within_rep_senate)
)

figure_F_results
```



#### SAVE ALL DATA IN RDATA FILE

```{r}

save(figure_A_results,
     figure_B_results,
     figure_C_results,
     figure_D_results,
     figure_E_results,
     figure_F_results,
     file = "paper_replication_data.rdata"
     )

```


# Part 2

## Bonacich Centrality and importance in bill sponsorships


### 108th Congress - SENATE

Obtain bonacich centrality measures of each senator in the 108th Congress.
```{r}
# Get the igraph object from the incidence matrix, but this time pull the weighted one so that we can extract who is the MAIN sponsor of each bill
graph_108 <- graph_from_incidence_matrix(
    incidence = senate_list$`108`$matrix,
    directed = FALSE,
    mode = "all",
    multiple = FALSE,
    weighted = NULL,
    add.names = NULL
)

# Get the bonacich power centralities of the nodes in the graph
centralities_108 <- as.data.frame(power_centrality(graph_108)[1:100])
colnames(centralities_108) <- c("power_centrality")

# Sort the centralities in descending order and extract the top ten nodes
top_ten_nodes <- centralities_108[order(-centralities_108$power_centrality), ][1:10]

# Subset the top ten- just to see who the most central senators are
top_ten_central_108_S <- subset(centralities_108, power_centrality %in% top_ten_nodes)
top_ten_central_108_S <- arrange(top_ten_central_108_S, desc(power_centrality))
top_ten_central_108_S
```

Now, we want to see if centrality can play a role in bill passage.

Extract weighted incidence matrix.
```{r}
# Extract incidence matrix for the current session, but this time weighted so the main sponsor is labeled as 2
senate_108_weighted <-
  incidence.from.congress(
    session = 108,
    format = "data",
    weighted = TRUE,
    types = c("s", "sjres")
  ) 
```

Extract the bills and incidence matrix.
```{r}
# Dataframe that shows the bills and the metadata pulled
# Main sponsor's entry is 2 in incidence matrix

bills_108 <- senate_108_weighted$bills
matrix_108 <- senate_108_weighted$matrix

# Print a few rows and columns
matrix_108[1:5, 1:5]
```

Extract the main sponsor of each bill.
```{r}
# Create an empty dataframe to store the results
result_df <- data.frame(Bill = character(), Senator = character(), stringsAsFactors = FALSE)

# Extract main sponsors
main_sponsors <- apply(matrix_108 == 2, 2, function(x) which(x == TRUE))

# Populate the dataframe with bill names and senator names
for (i in 1:length(main_sponsors)) {
  bill <- paste0("Bill_", i)
  sponsor <- paste0("Senator_", main_sponsors[[i]])
  senator_name <- row.names(matrix_108)[as.numeric(sub("Senator_", "", sponsor))]
  
  result_df <- rbind(result_df, data.frame(Bill = bill, Senator = senator_name))
}

# Remove the prefix text
result_df$Bill <- sub("Bill_", "", result_df$Bill)

# Set the index column of result_df
rownames(result_df) <- colnames(matrix_108[main_sponsors, ])

# Display the resulting dataframe
# results_df shows the main sponsor's name for each bill
print(result_df)

```
Add the main sponsor and their bonacich centrality to the bill metadata.
```{r}
# Add sponsor name corresponding to each bill into bills_108 (from results_df)
  # Join 'result_df' with 'bills_108' based on the 'bill' column
bills_108 <- merge(bills_108, result_df, by.x = "bill", by.y = 0, all.x = TRUE)
colnames(bills_108)[10:11] <- c("Bill index(from matrix)", "main_sponsor")

# Add sponsor's bonacich power centrality
  # Join 'centralities_108' with 'bills_108' based on the 'main_sponsor' column
bills_108 <- merge(bills_108, centralities_108, by.x = "main_sponsor", by.y = 0, all.x = TRUE)
colnames(bills_108)[12] <- "bonacich_sponsor"

# Create a new column 'status_code' based on 'status'
bills_108$status_code <- ifelse(bills_108$status == "Introduced", 0,
                                ifelse(bills_108$status == "Passed senate", 1,
                                       ifelse(bills_108$status == "Sent to president", 2, 
                                              ifelse(bills_108$status == "Became law", 3, NaN))))

# Display the modified dataframe
print(bills_108)
```

#### Question 1: Do senators with higher bonacich centrality get more bills passed (or get bills further through the process)?

*Note: It goes "Introduced" -> "Passed senate" -> "Sent to president" -> "Became law"*
```{r}
# Regression model
model1 <- lm(status_code ~ bonacich_sponsor, data = bills_108)
summary(model1)
```

This next result makes sense considering that the Republicans held the majority in the Senate in the 108th Congress.
```{r}
model2 <- lm(status_code ~ bonacich_sponsor + sponsor.party, data = bills_108)
summary(model2)
```


#### Question 2: Do senators with higher bonacich centrality get more senators from *opposing parties* to cosponsor their bills?

First add a column that shows how many cosponsoring senators were from the opposing party (different party than the main sponsor).
```{r}
# Add column "cosponsors_opposing_parties" that shows how many cosponsors of the opposing party the bill received
bills_108$cosponsors_opposing_parties <- ifelse(bills_108$sponsor.party == "R", bills_108$cosponsors.d + bills_108$cosponsors.i,
                                                ifelse(bills_108$sponsor.party == "D", bills_108$cosponsors.r + bills_108$cosponsors.i,
                                                       bills_108$cosponsors.r + bills_108$cosponsors.d))
# View df
bills_108
```

```{r}
# Regression model
model3 <- lm(cosponsors_opposing_parties ~ bonacich_sponsor, data = bills_108)
summary(model3)
```

```{r}
# Regression model
model4 <- lm(cosponsors_opposing_parties ~ bonacich_sponsor + sponsor.party, data = bills_108)
summary(model4)
```



### 108th Congress - HOUSE OF REPRESENTATIVES 

Obtain bonacich centrality measures of each representative in the 108th Congress.
```{r}
# Get the igraph object from the incidence matrix, but this time pull the weighted one so that we can extract who is the MAIN sponsor of each bill
graph_108_H <- graph_from_incidence_matrix(
    incidence = house_list$`108`$matrix,
    directed = FALSE,
    mode = "all",
    multiple = FALSE,
    weighted = NULL,
    add.names = NULL
)

# Get the bonacich power centralities of the nodes in the graph
centralities_108_H <- as.data.frame(power_centrality(graph_108_H, exponent = 0.9)[1:490])
colnames(centralities_108_H) <- c("power_centrality")

# Sort the centralities in descending order and extract the top ten nodes
top_ten_nodes_H <- centralities_108_H[order(-centralities_108_H$power_centrality), ][1:10]

# Subset the top ten- just to see who the most central senators are
top_ten_central_108_H <- subset(centralities_108_H, power_centrality %in% top_ten_nodes_H)
top_ten_central_108_H <- arrange(top_ten_central_108_H, desc(power_centrality))
top_ten_central_108_H
```
Now, we want to see if centrality can play a role in bill passage.

Extract weighted incidence matrix.
```{r}
# Extract incidence matrix for the current session, but this time weighted so the main sponsor is labeled as 2
house_108_weighted <-
  incidence.from.congress(
    session = 108,
    format = "data",
    weighted = TRUE,
    types = c("hr", "hjres")
  ) 
```

Extract the bills and incidence matrix.
```{r}
# Dataframe that shows the bills and the metadata pulled
# Main sponsor's entry is 2 in incidence matrix

bills_108_H <- house_108_weighted$bills
matrix_108_H <- house_108_weighted$matrix

# Print a few rows and columns
matrix_108_H[1:5, 1:5]
```

Extract the main sponsor of each bill.
```{r}
# Create an empty dataframe to store the results
result_df <- data.frame(Bill = character(), Representative = character(), stringsAsFactors = FALSE)

# Extract main sponsors
main_sponsors <- apply(matrix_108_H == 2, 2, function(x) which(x == TRUE))

# Populate the dataframe with bill names and senator names
for (i in 1:length(main_sponsors)) {
  bill <- paste0("Bill_", i)
  sponsor <- paste0("Rep_", main_sponsors[[i]])
  rep_name <- row.names(matrix_108_H)[as.numeric(sub("Rep_", "", sponsor))]
  
  result_df <- rbind(result_df, data.frame(Bill = bill, Representative = rep_name))
}

# Remove the prefix text
result_df$Bill <- sub("Bill_", "", result_df$Bill)

# Set the index column of result_df
rownames(result_df) <- colnames(matrix_108_H)

# Display the resulting dataframe
# results_df shows the main sponsor's name for each bill
print(result_df)

```

Add the main sponsor and their bonacich centrality to the bill metadata.
```{r}
# Add sponsor name corresponding to each bill into bills_108 (from results_df)
  # Join 'result_df' with 'bills_108' based on the 'bill' column
bills_108_H <- merge(bills_108_H, result_df, by.x = "bill", by.y = 0, all.x = TRUE)
colnames(bills_108_H)[10:11] <- c("Bill index(from matrix)", "main_sponsor")

# Add sponsor's bonacich power centrality
  # Join 'centralities_108' with 'bills_108' based on the 'main_sponsor' column
bills_108_H <- merge(bills_108_H, centralities_108_H, by.x = "main_sponsor", by.y = 0, all.x = TRUE)
colnames(bills_108_H)[12] <- "bonacich_sponsor"

# Create a new column 'status_code' based on 'status'
bills_108_H$status_code <- ifelse(bills_108_H$status == "Introduced", 0,
                                ifelse(bills_108_H$status == "Passed senate", 1,
                                       ifelse(bills_108_H$status == "Sent to president", 2, 
                                              ifelse(bills_108_H$status == "Became law", 3, NaN))))

# Display the modified dataframe
print(bills_108_H)
```

#### Question 1: Do representatives with higher bonacich centrality get more bills passed (or get bills further through the process)?

*Note: It goes "Introduced" -> "Passed senate" -> "Sent to president" -> "Became law"*
```{r}
# Regression model
model1_H <- lm(status_code ~ bonacich_sponsor, data = bills_108_H)
summary(model1_H)
```

This next result makes sense considering that the Republicans held the majority in the Senate in the 108th Congress.
```{r}
model2_H <- lm(status_code ~ bonacich_sponsor + sponsor.party, data = bills_108_H)
summary(model2_H)
```


#### Question 2: Do representatives with higher bonacich centrality get more senators from *opposing parties* to cosponsor their bills?

First add a column that shows how many cosponsoring senators were from the opposing party (different party than the main sponsor).
```{r}
# Add column "cosponsors_opposing_parties" that shows how many cosponsors of the opposing party the bill received
bills_108_H$cosponsors_opposing_parties <- ifelse(bills_108_H$sponsor.party == "R", bills_108_H$cosponsors.d + bills_108_H$cosponsors.i,
                                                ifelse(bills_108_H$sponsor.party == "D", bills_108_H$cosponsors.r + bills_108_H$cosponsors.i,
                                                       bills_108_H$cosponsors.r + bills_108_H$cosponsors.d))
# View df
bills_108_H
```

```{r}
# Regression model
model3_H <- lm(cosponsors_opposing_parties ~ bonacich_sponsor, data = bills_108_H)
summary(model3_H)
```

```{r}
# Regression model
model4_H <- lm(cosponsors_opposing_parties ~ bonacich_sponsor + sponsor.party, data = bills_108_H)
summary(model4_H)
```



## Senate - 116th (R) through 117th (D) TOGETHER

Obtain bonacich centrality measures of each senator in the 108th Congress.
```{r}
# Get the igraph object from the incidence matrix, but this time pull the weighted one so that we can extract who is the MAIN sponsor of each bill
graph_116 <- graph_from_incidence_matrix(
    incidence = senate_list$`116`$matrix,
    directed = FALSE,
    mode = "all",
    multiple = FALSE,
    weighted = NULL,
    add.names = NULL
)

# Get the igraph object from the incidence matrix, but this time pull the weighted one so that we can extract who is the MAIN sponsor of each bill
graph_117 <- graph_from_incidence_matrix(
    incidence = senate_list$`117`$matrix,
    directed = FALSE,
    mode = "all",
    multiple = FALSE,
    weighted = NULL,
    add.names = NULL
)

# Get the bonacich power centralities of the nodes in the graph
centralities_116 <- as.data.frame(power_centrality(graph_116)[1:100])
colnames(centralities_116) <- c("power_centrality")

centralities_117 <- as.data.frame(power_centrality(graph_117)[1:100])
colnames(centralities_117) <- c("power_centrality")
```


Now, we want to see if centrality can play a role in bill passage.

Extract weighted incidence matrix.
```{r}
# Extract incidence matrix for the current session, but this time weighted so the main sponsor is labeled as 2
senate_116_weighted <-
  incidence.from.congress(
    session = 116,
    format = "data",
    weighted = TRUE,
    types = c("s", "sjres")
  ) 

# Extract incidence matrix for the current session, but this time weighted so the main sponsor is labeled as 2
senate_117_weighted <-
  incidence.from.congress(
    session = 117,
    format = "data",
    weighted = TRUE,
    types = c("s", "sjres")
  ) 
```

Extract the bills and incidence matrix.
```{r}
# Dataframe that shows the bills and the metadata pulled
# Main sponsor's entry is 2 in incidence matrix

bills_116 <- senate_116_weighted$bills
matrix_116 <- senate_116_weighted$matrix

bills_117 <- senate_117_weighted$bills
matrix_117 <- senate_117_weighted$matrix

```

Extract the main sponsor of each bill.
```{r}
# Create an empty dataframe to store the results
result_df_116 <- data.frame(Bill = character(), Senator = character(), stringsAsFactors = FALSE)
result_df_117 <- data.frame(Bill = character(), Senator = character(), stringsAsFactors = FALSE)

# Extract main sponsors
main_sponsors_116 <- apply(matrix_116 == 2, 2, function(x) which(x == TRUE))

main_sponsors_117 <- apply(matrix_117 == 2, 2, function(x) which(x == TRUE))

# Populate the dataframe with bill names and senator names
for (i in 1:length(main_sponsors_116)) {
  bill <- paste0("Bill_", i)
  sponsor <- paste0("Senator_", main_sponsors_116[[i]])
  senator_name <- row.names(matrix_116)[as.numeric(sub("Senator_", "", sponsor))]
  
  result_df_116 <- rbind(result_df_116, data.frame(Bill = bill, Senator = senator_name))
}

for (i in 1:length(main_sponsors_117)) {
  bill <- paste0("Bill_", i)
  sponsor <- paste0("Senator_", main_sponsors_117[[i]])
  senator_name <- row.names(matrix_117)[as.numeric(sub("Senator_", "", sponsor))]
  
  result_df_117 <- rbind(result_df_117, data.frame(Bill = bill, Senator = senator_name))
}

# Remove the prefix text
result_df_116$Bill <- sub("Bill_", "", result_df_116$Bill)
result_df_117$Bill <- sub("Bill_", "", result_df_117$Bill)

# Set the index column of result_df
rownames(result_df_116) <- colnames(matrix_116)
rownames(result_df_117) <- colnames(matrix_117)

# Display the resulting dataframe
# results_df shows the main sponsor's name for each bill
print(result_df_116)
print(result_df_117)

```

Add the main sponsor and their bonacich centrality to the bill metadata.
```{r}
# Add sponsor name corresponding to each bill into bills_108 (from results_df)
  # Join 'result_df' with 'bills_108' based on the 'bill' column
bills_116 <- merge(bills_116, result_df_116, by.x = "bill", by.y = 0, all.x = TRUE)
colnames(bills_116)[10:11] <- c("Bill index(from matrix)", "main_sponsor")

bills_117 <- merge(bills_117, result_df_117, by.x = "bill", by.y = 0, all.x = TRUE)
colnames(bills_117)[10:11] <- c("Bill index(from matrix)", "main_sponsor")

# Add sponsor's bonacich power centrality
  # Join 'centralities_108' with 'bills_108' based on the 'main_sponsor' column
bills_116 <- merge(bills_116, centralities_116, by.x = "main_sponsor", by.y = 0, all.x = TRUE)
colnames(bills_116)[12] <- "bonacich_sponsor"

bills_117 <- merge(bills_117, centralities_117, by.x = "main_sponsor", by.y = 0, all.x = TRUE)
colnames(bills_117)[12] <- "bonacich_sponsor"

# Create a new column 'status_code' based on 'status'
bills_116$status_code <- ifelse(bills_116$status == "Introduced", 0,
                                ifelse(bills_116$status == "Passed senate", 1,
                                       ifelse(bills_116$status == "Sent to president", 2, 
                                              ifelse(bills_116$status == "Became law", 3, NaN))))

bills_117$status_code <- ifelse(bills_117$status == "Introduced", 0,
                                ifelse(bills_117$status == "Passed senate", 1,
                                       ifelse(bills_117$status == "Sent to president", 2, 
                                              ifelse(bills_117$status == "Became law", 3, NaN))))

# Display the modified dataframe
print(bills_116)
print(bills_117)
```

Merge 116 and 117.
```{r}
Senate_116_117 <- rbind(bills_116, bills_117)
```


#### Question 1: Do senators with higher bonacich centrality get more bills passed (or get bills further through the process)?

*Note: It goes "Introduced" -> "Passed senate" -> "Sent to president" -> "Became law"*
```{r}
# Regression model
model_116_117 <- lm(status_code ~ bonacich_sponsor, data = Senate_116_117)
summary(model_116_117)
```

#### Question 2: Do representatives with higher bonacich centrality get more senators from *opposing parties* to cosponsor their bills?

First add a column that shows how many cosponsoring senators were from the opposing party (different party than the main sponsor).
```{r}
# Add column "cosponsors_opposing_parties" that shows how many cosponsors of the opposing party the bill received
Senate_116_117$cosponsors_opposing_parties <- ifelse(Senate_116_117$sponsor.party == "R", Senate_116_117$cosponsors.d + Senate_116_117$cosponsors.i,
                                                ifelse(Senate_116_117$sponsor.party == "D", Senate_116_117$cosponsors.r + Senate_116_117$cosponsors.i,
                                                       Senate_116_117$cosponsors.r + Senate_116_117$cosponsors.d))
```

```{r}
# Regression model
model2_116_117 <- lm(cosponsors_opposing_parties ~ bonacich_sponsor, data = Senate_116_117)
summary(model2_116_117)
```
