---
title: "Networks Project"
output: html_document
date: "2023-06-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load Libraries.
```{r}
library(incidentally)     # install.packages("incidentally")
library(backbone)         # install.packages("backbone")
library(igraph)
library(tibble)
library(stringr)
```


## SENATE - Pull incidence matrices and data for each session 
```{r}
# Create an empty list to store the outputs
senate_incidences <- list()

# Define the range of sessions (years)
start_session <- 108
end_session <- 118

# Loop through each session and extract the incidence matrix
for (session in start_session:end_session) {
  # Initialize an empty variable to store the current session's output
  senate_session <- NULL
  
  tryCatch({
  # Extract incidence matrix for the current session
  senate_session <- incidence.from.congress(
    session = session,
    format = "data",
    types = c("s", "sjres")
  ) 
  
  # Generate a name for the output list item
  session_name <- paste0("senate_", session)
  
  # Save the incidence matrix to the output list
  senate_incidences[[session_name]] <- senate_session
  
  # Print a message indicating the progress
  cat("Session", session, "completed.\n")
  }, error = function(e) {
    # Print an error message if the extraction fails
    cat("Error extracting session", session, ": ", conditionMessage(e), "\n")
  })
}


# Print the names of the output list items
print(names(senate_incidences))

```



SENATE - Extract backbone using SDSM and save
```{r}

senate_list <- list()
  
for (i in seq_along(senate_incidences)) {
  tryCatch({
  # Save name to remember 
  name <- names(senate_incidences)[i]
    
  # First extract the matrix from the list extracted from incidentally
  object <- senate_incidences[[i]]
  
  # Extract backbone using the Stochastic Degree Sequence Model
  matrix <- as.matrix(sdsm(object$matrix, narrative = TRUE, signed = TRUE))
  
  # Then create the dataframe
  df <- createDataFrame(matrix)
  
  # Then create the legislators lists
  dems_list <- subset(object$legislator, party == "D")
  reps_list <- subset(object$legislator, party == "R")

  # Append to senate_list
  senate_list[[i]] <- tibble::lst(name, object, matrix, df, dems_list, reps_list)
  }, error = function(e) {
    # Print an error message if the extraction fails
    cat("Error extracting", name, ": ", conditionMessage(e), "\n")
  })
}

```



```{r}
names(senate_list) <- c(108, 111, 113, 114, 115, 116, 117, 118)
```

Save the output.
```{r}
save(senate_incidences, 
     senate_list, 
     file="senate_data.rdata")
```

*Note: The SDSM function is not working for sessions 109, 110, and 112*



## HOUSE - Pull incidence matrices and data for each session
```{r}
# Create an empty list to store the outputs
house_incidences <- list()

# Define the range of sessions (years)
start_session <- 108
end_session <- 118

# Loop through each session and extract the incidence matrix
for (session in start_session:end_session) {
  # Initialize an empty variable to store the current session's output
  house_session <- NULL
  
  tryCatch({
  # Extract incidence matrix for the current session
  house_session <- incidence.from.congress(
    session = session,
    format = "data",
    types = c("hr", "hjres")      # "hjres"    # "hr"
  )
  
  # Generate a name for the output list item
  session_name <- paste0("house_", session)
  
  # Save the incidence matrix to the output list
  house_incidences[[session_name]] <- house_session
  
  # Print a message indicating the progress
  cat("Session", session, "completed.\n")
  }, error = function(e) {
    # Print an error message if the extraction fails
    cat("Error extracting session", session, ": ", conditionMessage(e), "\n")
  })
}


# Print the names of the output list items
print(names(house_incidences))

```


HOUSE - Extract backbone using SDSM and save
```{r}
house_list <- list()
  
for (i in seq_along(house_incidences)) {
  # Save name to remember 
  name <- names(house_incidences)[i]
    
  # First extract the matrix from the list extracted from incidentally
  object <- house_incidences[[i]]
  
  # Extract backbone using the Stochastic Degree Sequence Model
  matrix <- as.matrix(sdsm(object$matrix, narrative = TRUE, signed = TRUE))
  
  # Then create the dataframe
  df <- createDataFrame(matrix)
  
  # Then create the legislators lists
  dems_list <- subset(object$legislator, party == "D")
  reps_list <- subset(object$legislator, party == "R")

  # Append to house_list
  house_list[[i]] <- tibble::lst(name, object, matrix, df, dems_list, reps_list)
}

```

```{r}
names(house_list) <- seq(108, 118)
```

Save the output.
```{r}
save(house_incidences,
     house_list,
     file="house_data.rdata")
```

*Note: The incidence function was failing to pull the House resolution ("hr") bills so the data for House 116th session is only based on 388 bills that were house joint resolution ("hjres") bills*



# Recreating the figures

Function to extract pairs and their tie
```{r}
## Function 
createDataFrame <- function(matrix) {
  # Define upper triangle values
  upper_triangle <- upper.tri(matrix)
  upper_triangle_values <- matrix[upper_triangle]
  
  # Get the row and column names of the upper triangle elements
  rep_pairs <- expand.grid(colnames(matrix), colnames(matrix))
  rep_pairs <- rep_pairs[upper_triangle, ]
  
  # Create the dataframe with representative pairs
  df <- data.frame(rep1 = rep_pairs$Var1,
                   rep2 = rep_pairs$Var2,
                   tie = upper_triangle_values)
  
  # Return the df
  return(df)
  
  # Remove unnecessary stuff
  rm(upper_triangle, upper_triangle_values, rep_pairs, df)
}
```


### FIGURE A - HOUSE
```{r}

prop_pos_all_list_house <- numeric(length(house_list))
prop_neg_all_list_house <- numeric(length(house_list))

for (i in seq_along(house_list)) {
  matrix <- house_list[[i]]$matrix
  prop_pos_all_list_house[i] <- nrow(subset(createDataFrame(matrix), tie == 1)) / nrow(createDataFrame(matrix))
  prop_neg_all_list_house[i] <- nrow(subset(createDataFrame(matrix), tie == -1)) / nrow(createDataFrame(matrix))
}

```

```{r}
# Create df 
figure_A_results <- data.frame (
  session = names(house_list),
  prop_pos_all_H = c(prop_pos_all_list_house),
  prop_neg_all_H = c(prop_neg_all_list_house)
)

figure_A_results
```

#### FIGURE B - SENATE


```{r}

prop_pos_all_list_sen <- numeric(length(senate_list))
prop_neg_all_list_sen<- numeric(length(senate_list))

for (i in seq_along(senate_list)) {
  matrix <- senate_list[[i]]$matrix
  prop_pos_all_list_sen[i] <- nrow(subset(createDataFrame(matrix), tie == 1)) / nrow(createDataFrame(matrix))
  prop_neg_all_list_sen[i] <- nrow(subset(createDataFrame(matrix), tie == -1)) / nrow(createDataFrame(matrix))
}


```

```{r}
# Create df 
figure_B_results <- data.frame (
  session = names(senate_list),
  prop_pos_all_S = c(prop_pos_all_list_sen),
  prop_neg_all_S = c(prop_neg_all_list_sen)
)

figure_B_results
```


```{r}
df<-house_list$`108`$df

nrow(df) == nrow(subset(df, tie==-1)) + nrow(subset(df, tie==1)) + nrow(subset(df, tie==0))
```


### FIGURE C - HOUSE

```{r} 

prop_pos_within_house <- numeric(length(house_list))
prop_neg_between_house <- numeric(length(house_list))

for (i in seq_along(house_list)) {
  # Then create the dataframe
  df <- createDataFrame(house_list[[i]]$matrix)
  
  # Then create the legislators lists
  dems_list <- house_list[[i]]$dems_list
  reps_list <- house_list[[i]]$reps_list
  
  # Add columns "rep1_party" and "rep2_party"
  df$rep1_party <- ifelse(df$rep1 %in% dems_list$name, "D", ifelse(df$rep1 %in% reps_list$name, "R", "NA"))
  df$rep2_party <- ifelse(df$rep2 %in% dems_list$name, "D", ifelse(df$rep2 %in% reps_list$name, "R", "NA"))

  # Append proportions
  prop_pos_within_house[i] <- nrow(subset((subset(df, rep1_party == rep2_party & rep1_party != "NA")), tie == 1)) / nrow(subset(df, tie == 1))
  
  prop_neg_between_house[i] <- nrow(subset((subset(df, rep1_party != rep2_party & rep1_party != "NA" & rep2_party != "NA")), tie == -1)) / nrow(subset(df, tie == -1))
}

```

```{r}
# Create df 
figure_C_results <- data.frame (
  session = names(house_list),
  prop_pos_within_H = c(prop_pos_within_house),
  prop_neg_between_H = c(prop_neg_between_house)
)

figure_C_results
```

#### FIGURE D - SENATE
```{r} 

prop_pos_within_senate <- numeric(length(senate_list))
prop_neg_between_senate <- numeric(length(senate_list))

for (i in seq_along(senate_list)) {
  df <- createDataFrame(senate_list[[i]]$matrix)
  
  # Then create the legislators lists
  dems_list <- senate_list[[i]]$dems_list
  reps_list <- senate_list[[i]]$reps_list
    
  # Add columns "rep1_party" and "rep2_party"
  df$rep1_party <- ifelse(df$rep1 %in% dems_list$name, "D", ifelse(df$rep1 %in% reps_list$name, "R", "NA"))
  df$rep2_party <- ifelse(df$rep2 %in% dems_list$name, "D", ifelse(df$rep2 %in% reps_list$name, "R", "NA"))

  # Append proportions
  prop_pos_within_senate[i] <- nrow(subset((subset(df, rep1_party == rep2_party & rep1_party != "NA")), tie == 1)) / nrow(subset(df, tie == 1))
    
  prop_neg_between_senate[i] <- nrow(subset((subset(df, rep1_party != rep2_party & rep1_party != "NA" & rep2_party != "NA")), tie == -1)) / nrow(subset(df, tie == -1))
}

```

```{r}
# Create df 
figure_D_results <- data.frame (
  session = names(senate_list),
  prop_pos_within_S = c(prop_pos_within_senate),
  prop_neg_between_S = c(prop_neg_between_senate)
)

figure_D_results
```


#### FIGURE E - HOUSE

Get the proportion of negative ties for each party
```{r}

prop_neg_within_dem_house <- numeric(length(house_list))
prop_neg_within_rep_house <- numeric(length(house_list))
  

for (i in seq_along(house_list)) {
  df <- createDataFrame(house_list[[i]]$matrix)
  
  # Then extract the legislators lists
  dems_list <- house_list[[i]]$dems_list
  reps_list <- house_list[[i]]$reps_list
  
  # Add columns "rep1_party" and "rep2_party"
  df$rep1_party <- ifelse(df$rep1 %in% dems_list$name, "D", ifelse(df$rep1 %in% reps_list$name, "R", "NA"))
  df$rep2_party <- ifelse(df$rep2 %in% dems_list$name, "D", ifelse(df$rep2 %in% reps_list$name, "R", "NA"))
  
  # Append proportions
  prop_neg_within_dem_house[i] <- nrow(subset((subset(df, rep1_party == rep2_party & rep1_party == "D")), tie == -1)) / nrow(subset(df, rep1_party == rep2_party & rep1_party == "D"))
    
  prop_neg_within_rep_house[i] <- nrow(subset((subset(df, rep1_party == rep2_party & rep1_party == "R")), tie == -1)) / nrow(subset(df, rep1_party == rep2_party & rep1_party == "R"))
    
}

```

```{r}
# Create df 
figure_E_results <- data.frame (
  session = names(house_list),
  prop_neg_within_dem_H = c(prop_neg_within_dem_house),
  prop_neg_within_rep_H = c(prop_neg_within_rep_house)
)

figure_E_results
```




#### FIGURE F - SENATE

Get the proportion of negative ties for each party
```{r}

prop_neg_within_dem_senate <- numeric(length(senate_list))
prop_neg_within_rep_senate <- numeric(length(senate_list))
  

for (i in seq_along(senate_list)) {
  df <- createDataFrame(senate_list[[i]]$matrix)
  
  # Then extract the legislators lists
  dems_list <- senate_list[[i]]$dems_list
  reps_list <- senate_list[[i]]$reps_list
  
  # Add columns "rep1_party" and "rep2_party"
  df$rep1_party <- ifelse(df$rep1 %in% dems_list$name, "D", ifelse(df$rep1 %in% reps_list$name, "R", "NA"))
  df$rep2_party <- ifelse(df$rep2 %in% dems_list$name, "D", ifelse(df$rep2 %in% reps_list$name, "R", "NA"))
  
  # Append proportions
  prop_neg_within_dem_senate[i] <- nrow(subset((subset(df, rep1_party == rep2_party & rep1_party == "D")), tie == -1)) / nrow(subset(df, rep1_party == rep2_party & rep1_party == "D"))
    
  prop_neg_within_rep_senate[i] <- nrow(subset((subset(df, rep1_party == rep2_party & rep1_party == "R")), tie == -1)) / nrow(subset(df, rep1_party == rep2_party & rep1_party == "R"))
    
}

```

```{r}
# Create df 
figure_F_results <- data.frame (
  session = names(senate_list),
  prop_neg_within_dem_S = c(prop_neg_within_dem_senate),
  prop_neg_within_rep_S = c(prop_neg_within_rep_senate)
)

figure_F_results
```



#### SAVE ALL DATA IN RDATA FILE

```{r}

save(figure_A_results,
     figure_B_results,
     figure_C_results,
     figure_D_results,
     figure_E_results,
     figure_F_results,
     file = "paper_replication_data.rdata"
     )

```
