---
title: "Networks_project_votes"
author: "Lenka Sefcakova, Antonia George"
date: "2023-06-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Co-voting patterns on congress sitting analysis

Data used:

-   `HSall_votes` :

    -   **congress**: Integer `1+`. The number of the congress that this member's row refers to. e.g. `115` for the 115th Congress (2017-2019)

    -   **chamber**: `House`, `Senate`, or `President`. The chamber in which the member served.

    -   **rollnumber**: Integer `1+`. Starts from 1 in the first rollcall of each congress. Excludes quorum calls and vacated votes.

    -   **icpsr**: Integer `1-99999`. This is an ID code which identifies the member in question. In general, each member receives a single ICPSR identifier applicable to their entire career. A small number of members have received more than one: this can occur for members who have switched parties; as well as members who subsequently become president. Creating a new identifier allows a new NOMINATE estimate to be produced for separate appearances of a member in different roles.

    -   **cast_code**: Integer `0-9`. Indicator of how the member voted.

    -   **prob**: Estimated probability, based on NOMINATE, of the member making the vote as recorded.

-   `HSall_members` (use icpsr and party_code)

    -   **congress**: Integer `1+`. The number of the congress that this member's row refers to. e.g. `115` for the 115th Congress (2017-2019)

    -   **chamber**: `House`, `Senate`, or `President`. The chamber in which the member served.

    -   **icpsr**: Integer `1-99999`. This is an ID code which identifies the member in question. In general, each member receives a single ICPSR identifier applicable to their entire career. A small number of members have received more than one: this can occur for members who have switched parties; as well as members who subsequently become president. Creating a new identifier allows a new NOMINATE estimate to be produced for separate appearances of a member in different roles.

    -   **state_icpsr**: Integer `0-99`. Identifier for the state represented by the member.

    -   **district_code**: Integer `0-99`. Identifier for the district that the member represents within their state (e.g. `3` for the Alabama 3rd Congressional District). Senate members are given `district_code` 0. Members who represent historical "at-large" districts are assigned `99`, `98`, or `1` in various circumstances.

    -   **state_abbrev**: String. Two-character postal abbreviation for state (e.g. `MO` for Missouri).

    -   **party_code**: Integer `1-9999`. Identifying code for the member's party. Please see [documentation for Party Data](https://voteview.com/articles/data_help_parties) for more information about which `party_code` identifiers refer to which parties.

    -   **occupancy**: Integer `1+`. ICPSR occupancy code. This item is considered legacy or incomplete information and has not been verified. In general, members receive `0` if they are the only occupant, `1` if they are the first occupant, `2` if they are the second occupant, etc.

    -   **last_means**: Integer `1-5`. ICPSR Attain-Office Code. This is an indicator that reflects the member's last means of attaining office. This item is considered legacy or incomplete information and has not been verified. Members received `1` if they were elected in a general election, `2` if elected by special election, `3` if directly elected by a state legislature, and `5` if appointed.

    -   **bioname**: String. Name of the member, surname first. For most members, agrees with the Biographical Directory of Congress.

    -   **bioguide_id**: String. Member identifier in the Biographical Directory of Congress.

    -   **born**: Integer. Year of member's birth.

    -   **died**: Integer. Year of member's death.

## Making a network of co-votes per one congress.

```{r}
library(igraph)
library(ggplot2)
library(dplyr)
```

```{r}
data_votes <- read.csv('congress_data/HSall_votes.csv')
```

```{r}
data_members_all <- read.csv('congress_data/HSall_members.csv')
```

```{r}
data_members <- data_members_all[c('icpsr','congress','chamber', 'party_code')]
```

To do: (data_votes)

1.  subset data based on congress and chamber
2.  replace cast_code with simplified version 1,..9, -\> 1,6,9
3.  map party code from data_members
4.  build adjacency matrix with weighted edges based on co-voting (build edges based on roll number, add them over on congress and house)

```{r}
library(ggplot2)
congress_ids <-unique(data_votes$congress)
paste('The covered congresses in the data are ( from / to ) :')
length(congress_ids)
min(congress_ids)
max(congress_ids)
paste('The covered chambers in the data are : ')
congress_chambers <-unique(data_votes$chamber)
congress_chambers


```

```{r}
merged_data <- merge(data_votes, data_members)
```

```{r}
congress_ids_merge <-unique(merged_data$congress)
paste('The covered congresses in the data are ( from / to ) :')
length(congress_ids_merge)
min(congress_ids_merge)
max(congress_ids_merge)
paste('The covered chambers in the data are : ')
congress_chambers_merge <-unique(merged_data$chamber)
congress_chambers_merge
```

These ids below are the presidents and an anonymous person sitting in the first congress which we will discard as we do not have information about their initial party beliefs.

```{r}
cast_code_map <- function(vec) {
  mapping <- c("0", "1", "1", "1", "6", "6", "6", "9", "9","9")
  vec[vec %in% c(2,3)] = 1
  vec[vec %in% c(4,5)] = 6
  vec[vec %in% c(7,8)] = 9
  return(vec)
}

cast_code_bool <- function(vec) {
  mapping <- c('0','1','1','1','0','0','0','0','0','0')
  vec <- cast_code_map(vec)
  vec[vec %in% c(6,9,0)] = 0
  return(vec)
}
```

```{r}
library(backbone)

build_adj_mat <- function(congress_id, chamber_id) {
  sub <- merged_data[(merged_data$congress == congress_id) ,]
  sub <- sub[(sub$chamber == chamber_id) ,]
  sub$cast_code <- cast_code_bool(sub$cast_code)
  sub <- sub[(sub$cast_code == 1) ,]
  make_g <- as.data.frame(sub$icpsr)
  colnames(make_g) <- c('icpsr')
  colnames(make_g) <- as.character(colnames(make_g))
  make_g$rollnumber <- sub$rollnumber
  

  B <- as_adjacency_matrix(graph.data.frame(make_g, directed = FALSE))
  
  column_names <- unique(make_g$rollnumber)
  column_ind <- match(column_names, colnames(B))

  B <- B[-column_ind, column_ind, drop = FALSE]

  bbone <- sdsm(as.matrix(B), narrative = FALSE, signed = TRUE)
  
  return(bbone)
}
```

```{r}
result <- data_members %>%
  filter(chamber != "President") %>%
  group_by(congress, chamber,party_code) %>%
  summarise(num_entries = n()) %>%
  arrange(congress, chamber,party_code)%>%
  group_by(congress, chamber) %>%
  top_n(1, num_entries) %>%
  ungroup()
```

```{r}

```

```{r}
createDataFrame <- function(matrix) {
  # Define upper triangle values
  upper_triangle <- upper.tri(matrix)
  upper_triangle_values <- matrix[upper_triangle]
  
  # Get the row and column names of the upper triangle elements
  rep_pairs <- expand.grid(colnames(matrix), colnames(matrix))
  rep_pairs <- rep_pairs[upper_triangle, ]
  
  # Create the dataframe with representative pairs
  df <- data.frame(rep1 = rep_pairs$Var1,
                   rep2 = rep_pairs$Var2,
                   tie = upper_triangle_values)
  
  # Return the df
  return(df)
  
  # Remove unnecessary stuff
  rm(upper_triangle, upper_triangle_values, rep_pairs, df)
}
```

```{r}
senate_mat_list <- lapply(congress_ids,function(mat) build_adj_mat(mat,'Senate'))
house_mat_list <- lapply(congress_ids,function(mat) build_adj_mat(mat,'House'))
```

```{r}
senate_pairs <- lapply(senate_mat_list,createDataFrame)
house_pairs <- lapply(house_mat_list,createDataFrame)
```

```{r}
prop_pos_all_list_house <- numeric(length(house_pairs))
prop_neg_all_list_house <- numeric(length(house_pairs))

prop_pos_all_list_senate <- numeric(length(senate_pairs))
prop_neg_all_list_senate <- numeric(length(senate_pairs))

for (i in seq_along(house_pairs)) {
  df <- house_pairs[[i]]
  prop_pos_all_list_house[i] <- nrow(subset(df, tie == 1)) / nrow(df)
  prop_neg_all_list_house[i] <- nrow(subset(df, tie == -1)) / nrow(df)
  
  df <- senate_pairs[[i]]
  prop_pos_all_list_senate[i] <- nrow(subset(df, tie == 1)) / nrow(df)
  prop_neg_all_list_senate[i] <- nrow(subset(df, tie == -1)) / nrow(df)
}

figure_A_results <- data.frame (
  congress = congress_ids,
  prop_pos_all = c(prop_pos_all_list_house),
  prop_neg_all = c(prop_neg_all_list_house)
)

figure_B_results <- data.frame (
  congress = congress_ids,
  prop_pos_all = c(prop_pos_all_list_senate),
  prop_neg_all = c(prop_neg_all_list_senate)
)

```

```{r}
power_senate <- result[result$chamber == 'Senate',]
power_senate<- power_senate[power_senate$party_code == '200',]
power_house <- result[result$chamber == 'House',]
power_house<- power_house[power_house$party_code == '200',]
```

```{r}
library(ggplot2)
plot_A <- ggplot(figure_A_results, aes(x = congress)) +
  geom_line(aes(y = prop_pos_all, color = "Positive")) +
  geom_line(aes(y = prop_neg_all, color = "Negative")) +
  labs(x = "Congress", y = "Proportion", color = "Line") +
  scale_color_manual(values = c("Positive" = "blue", "Negative" = "red"))+ggtitle("Figure A - House") + # Set y-axis range
    geom_rect(
    data = power_house,
    aes(xmin = congress - 0.5, xmax = congress + 0.5,ymin = 0,ymax = 1),
    fill = "darkred",
    alpha = 0.2,
    inherit.aes = FALSE)+scale_x_continuous(labels = function(x) as.integer(x))

plot_B <- ggplot(figure_B_results, aes(x = congress)) +
  geom_line(aes(y = prop_pos_all, color = "Positive")) +
  geom_line(aes(y = prop_neg_all, color = "Negative")) +
  labs(x = "Congress", y = "Proportion", color = "Line") +
  scale_color_manual(values = c("Positive" = "blue", "Negative" = "red"))+ggtitle("Figure B - Senate") + # Set y-axis range
    geom_rect(
    data = power_senate,
    aes(xmin = congress - 0.5, xmax = congress + 0.5,ymin = 0,ymax = 1),
    fill = "darkred",
    alpha = 0.2,
    inherit.aes = FALSE)+scale_x_continuous(labels = function(x) as.integer(x))

plots_A_B <- grid.arrange(plot_A, plot_B, ncol = 2,widths = c(10,10),heights = c(3,3))


```

```{r}
prop_pos_within_house <- numeric(length(house_pairs))
prop_neg_between_house <- numeric(length(house_pairs))

prop_pos_within_senate <- numeric(length(senate_pairs))
prop_neg_between_senate <- numeric(length(senate_pairs))
```

data_members_all

```{r}
for (i in seq_along(house_pairs)) {

  df <- house_pairs[[i]]
  
  data_members <- filter(data_members_all, congress==i & chamber == 'House')
  merged_df <- merge(df, data_members, by.x = "rep1", by.y = "icpsr", all.x = TRUE, all.y = FALSE)
  merged_df <- merge(merged_df, data_members, by.x = "rep2", by.y = "icpsr", all.x = TRUE, all.y = FALSE, suffixes = c("_rep1", "_rep2"))
  
  result_df <- merged_df[, c("rep1", "rep2", "tie", "party_code_rep1", "party_code_rep2")]
  result_df$same_party_bool <- ifelse(result_df$party_code_rep1 == result_df$party_code_rep2, 1, 0)

  prop_pos_within_house[[i]] <- nrow(subset((subset(result_df, same_party_bool ==1)), tie == 1)) / nrow(subset(result_df, same_party_bool ==1))
  
  prop_neg_between_house[[i]] <- nrow(subset((subset(result_df, same_party_bool ==0)), tie == -1)) / nrow(subset(result_df, same_party_bool ==0))
  
}

for (i in seq_along(senate_pairs)) {
  df <- senate_pairs[[i]]
  
  data_members <- filter(data_members_all, congress == i & chamber == 'Senate')
  merged_df <- merge(df, data_members, by.x = "rep1", by.y = "icpsr", all.x = TRUE, all.y = FALSE)
  merged_df <- merge(merged_df, data_members, by.x = "rep2", by.y = "icpsr", all.x = TRUE, all.y = FALSE, suffixes = c("_rep1", "_rep2"))
  
  result_df <- merged_df[, c("rep1", "rep2", "tie", "party_code_rep1", "party_code_rep2")]
  result_df$same_party_bool <- ifelse(result_df$party_code_rep1 == result_df$party_code_rep2, 1, 0)
  
  prop_pos_within_senate[[i]] <- nrow(subset((subset(result_df, same_party_bool == 1)), tie == 1)) / nrow(subset(result_df, same_party_bool == 1))
  
  prop_neg_between_senate[[i]] <- nrow(subset((subset(result_df, same_party_bool == 0)), tie == -1)) / nrow(subset(result_df, same_party_bool == 0))
}

```

```{r}

figure_C_results <- data.frame (
  congress = congress_ids,
  prop_pos_all = c(prop_pos_within_house),
  prop_neg_all = c(prop_neg_between_house)
)

figure_D_results <- data.frame (
  congress = congress_ids,
  prop_pos_all = c(prop_pos_within_senate),
  prop_neg_all = c(prop_neg_between_senate)
)

```

```{r}
library(ggplot2)

plot_A <-ggplot(figure_A_results, aes(x = congress)) +
  geom_line(aes(y = prop_pos_all, color = "Positive")) +
  geom_line(aes(y = prop_neg_all, color = "Negative")) +
  labs(x = "Congress", y = "Proportion", color = "Line") +
  scale_color_manual(values = c("Positive" = "blue", "Negative" = "red"))+ggtitle("Figure A - House") + # Set y-axis range
    geom_rect(
    data = power_house,
    aes(xmin = congress - 0.5, xmax = congress + 0.5,ymin = 0,ymax = 1),
    fill = "darkred",
    alpha = 0.2,
    inherit.aes = FALSE)+scale_x_continuous(labels = function(x) as.integer(x))

plot_B <- ggplot(figure_B_results, aes(x = congress)) +
  geom_line(aes(y = prop_pos_all, color = "Positive")) +
  geom_line(aes(y = prop_neg_all, color = "Negative")) +
  labs(x = "Congress", y = "Proportion", color = "Line") +
  scale_color_manual(values = c("Positive" = "blue", "Negative" = "red"))+ggtitle("Figure B - Senate")  + # Set y-axis range
    geom_rect(
    data = power_senate,
    aes(xmin = congress - 0.5, xmax = congress + 0.5,ymin = 0,ymax = 1),
    fill = "darkred",
    alpha = 0.2,
    inherit.aes = FALSE)+scale_x_continuous(labels = function(x) as.integer(x))
plots_A_B <- grid.arrange(plot_A, plot_B, ncol = 2,widths = c(10,10),heights = c(3,3))

plot_C <- ggplot(figure_C_results, aes(x = congress)) +
  geom_line(aes(y = prop_pos_all, color = "Positive")) +
  geom_line(aes(y = prop_neg_all, color = "Negative")) +
  labs(x = "Congress", y = "Proportion", color = "Line") +
  scale_color_manual(values = c("Positive" = "blue", "Negative" = "red"))+ggtitle("Figure C - House") + # Set y-axis range
    geom_rect(
    data = power_house,
    aes(xmin = congress - 0.5, xmax = congress + 0.5,ymin = 0,ymax = 1),
    fill = "darkred",
    alpha = 0.2,
    inherit.aes = FALSE)+scale_x_continuous(labels = function(x) as.integer(x))

plot_D <- ggplot(figure_D_results, aes(x = congress)) +
  geom_line(aes(y = prop_pos_all, color = "Positive")) +
  geom_line(aes(y = prop_neg_all, color = "Negative")) +
  labs(x = "Congress", y = "Proportion", color = "Line") +
  scale_color_manual(values = c("Positive" = "blue", "Negative" = "red"))+ggtitle("Figure D - Senate")  + # Set y-axis range
    geom_rect(
    data = power_senate,
    aes(xmin = congress - 0.5, xmax = congress + 0.5,ymin = 0,ymax = 1),
    fill = "darkred",
    alpha = 0.2,
    inherit.aes = FALSE)+scale_x_continuous(labels = function(x) as.integer(x))

plots_C_D <- grid.arrange(plot_C, plot_D, ncol = 2,widths = c(10,10),heights = c(3,3))

ggsave(filename = "figures/A_&_B_votes.jpg", plot = plots_A_B, device = "jpeg")
ggsave(filename = "figures/C_&_D_votes.jpg", plot = plots_C_D, device = "jpeg")

```

```{r}

library(ggplot2)

plot_A <-ggplot(figure_A_results, aes(x = congress)) +
  geom_line(aes(y = prop_pos_all, color = "Positive")) +
  geom_line(aes(y = prop_neg_all, color = "Negative")) +
  labs(x = "Congress", y = "Proportion", color = "Line") +
  scale_color_manual(values = c("Positive" = "blue", "Negative" = "red"))+ggtitle("Figure A - House") + # Set y-axis range
    geom_rect(
    data = power_house,
    aes(xmin = congress - 0.5, xmax = congress + 0.5,ymin = 0,ymax = 1),
    fill = "darkred",
    alpha = 0.2,
    inherit.aes = FALSE)+scale_x_continuous(labels = function(x) as.integer(x), limits = c(108,118))

plot_B <- ggplot(figure_B_results, aes(x = congress)) +
  geom_line(aes(y = prop_pos_all, color = "Positive")) +
  geom_line(aes(y = prop_neg_all, color = "Negative")) +
  labs(x = "Congress", y = "Proportion", color = "Line") +
  scale_color_manual(values = c("Positive" = "blue", "Negative" = "red"))+ggtitle("Figure B - Senate")  + # Set y-axis range
    geom_rect(
    data = power_senate,
    aes(xmin = congress - 0.5, xmax = congress + 0.5,ymin = 0,ymax = 1),
    fill = "darkred",
    alpha = 0.2,
    inherit.aes = FALSE)+scale_x_continuous(labels = function(x) as.integer(x), limits = c(108,118))
plots_A_B <- grid.arrange(plot_A, plot_B, ncol = 2,widths = c(10,10),heights = c(3,3))

plot_C <- ggplot(figure_C_results, aes(x = congress)) +
  geom_line(aes(y = prop_pos_all, color = "Positive")) +
  geom_line(aes(y = prop_neg_all, color = "Negative")) +
  labs(x = "Congress", y = "Proportion", color = "Line") +
  scale_color_manual(values = c("Positive" = "blue", "Negative" = "red"))+ggtitle("Figure C - House") + # Set y-axis range
    geom_rect(
    data = power_house,
    aes(xmin = congress - 0.5, xmax = congress + 0.5,ymin = 0,ymax = 1),
    fill = "darkred",
    alpha = 0.2,
    inherit.aes = FALSE)+scale_x_continuous(labels = function(x) as.integer(x), limits = c(108,118))

plot_D <- ggplot(figure_D_results, aes(x = congress)) +
  geom_line(aes(y = prop_pos_all, color = "Positive")) +
  geom_line(aes(y = prop_neg_all, color = "Negative")) +
  labs(x = "Congress", y = "Proportion", color = "Line") +
  scale_color_manual(values = c("Positive" = "blue", "Negative" = "red"))+ggtitle("Figure D - Senate")  + # Set y-axis range
    geom_rect(
    data = power_senate,
    aes(xmin = congress - 0.5, xmax = congress + 0.5,ymin = 0,ymax = 1),
    fill = "darkred",
    alpha = 0.2,
    inherit.aes = FALSE)+scale_x_continuous(labels = function(x) as.integer(x), limits = c(108,118))

plots_C_D <- grid.arrange(plot_C, plot_D, ncol = 2,widths = c(10,10),heights = c(3,3))

ggsave(filename = "figures/A_&_B_votes_current.jpg", plot = plots_A_B, device = "jpeg")
ggsave(filename = "figures/C_&_D_votes_current.jpg", plot = plots_C_D, device = "jpeg")
```
